---
title: "Data preparation and Objectives"
author: "Benzon Carlitos Salazar"
date: "1/8/2022"
output: pdf_document
---

```{r setup, message = FALSE, warning = FALSE, include = FALSE}
library(magrittr)

red_wine_dataset <- read.csv(here::here("data-raw", "winequality-red.csv"), header = TRUE, sep = ";")
```

# Save to R dataframe
```{r clean_and_save}
red_col_list <-
  colnames(red_wine_dataset[1:11]) %<>%
  gsub("[[:punct:]]", "_", .) %>%
  gsub("[[:space:]]", "_", .)

renamed_red_wine <-
  paste0(tolower(red_col_list))
  
red_wine_dataset %>%
  dplyr::rename_at(1:11, ~renamed_red_wine) %>%
  saveRDS(here::here("data", "red_wine_dataset.rds"))
```

# Objectives
We want to make a predictive model on the quality of **red wine** based on chemical properties of wine.

The following analytical approaches are taken:

* Multiple regression: the *Quality* response is assumed to be a continuous variable and is predicted by the independent 
predictors, all of which are continuous.
* Regression Tree
* Classification of wines based on the chemical properties: Unsupervised analysis
* Classification of wines based on the chemical properties: K-nearest neighbors classification

```{r clean_up, include = FALSE}
rm(list = ls(all.names = TRUE)) # clear all objects including hidden objects
invisible(gc()) # free up memory
```